---
title: "project!!!!!"
output: html_notebook
---

```{r link cloned repo so you can pull/commit/push}
# RUN THESE ALL ONCE!!! un-comment the code blocks as you go
#install.packages(c("usethis", "gitcreds"))

# then run this:
#usethis::create_github_token()

# which opens GitHub in your browser, generate a token
# COPY it before closing the page

# run this and paste token into terminal when prompted
#gitcreds::gitcreds_set()

# done! :) make sure to re-comment the code blocks
```

```{r libaries, include=FALSE}
# make sure to download any you dont have with: install.packages("package_name")
library(here)
library(corrplot)
library(tidyverse)
```

# Question 1 : Exploratory Analysis and Model Specification

```{r prep}
Demographic <- read.table(here("data", "Demographic.txt")) # more robust read-in

# change column names to match correct variables
colnames(Demographic) <- c('ID','County','state','Land_area','total_population',
                          'Population_18to34','Population_65','Physicians','beds',
                          'Y_i','Graduate_highschool','Graduate_Bachelor','below_poverty',
                          'unemployment','capita_income','personal_income','Geographic_region')
Demographic2 <- Demographic

# Variable names and dimensions of the table
head(Demographic2)
names(Demographic2)
dim(Demographic2)
summary(Demographic2)

Demographic2$Y_i <- log(Demographic$Y_i) # log transform Y

colnames(Demographic2)[10] <- "LogY_i"

region1 <- Demographic2[Demographic2$Geographic_region == 1,]
region2 <- Demographic2[Demographic2$Geographic_region == 2,]
region3 <- Demographic2[Demographic2$Geographic_region == 3,]
region4 <- Demographic2[Demographic2$Geographic_region == 4,]

n1<-nrow(region1)
n2<-nrow(region2)
n3<-nrow(region3)
n4<-nrow(region4)
```

```{r corr plots, warning=FALSE}
# general corr
cor(Demographic2[,4:17])
corrplot(cor(Demographic2[,4:17]))

# per region corr
regions <- list(region1, region2, region3, region4)
top_list <- list()

for (i in seq_along(regions)) {
  
  cat("FOR REGION: ", i, "\n")
  
  region <- regions[[i]]
  cors <- sapply(region[, 4:17], function(x) cor(x, region$LogY_i))
  ranked <- sort(abs(cors), decreasing = TRUE)
  
  top_list[[i]] <- head(ranked, 5)
  
  corrplot(cor(region$LogY_i, region[,4:17]))
  
  print(top_list[[i]])
}
```
## Region 1

```{r}
# log(Y_i) = Beta_0 + Beta_1*total_population

boxplot(cbind(region1$LogY_i, region1$total_population),names = c("Total serious crimes (log-scale)", "Total Population"))

pairs(cbind(region1$LogY_i, region1$total_population),labels=c("Total serious crimes  (log-scale)", "Total Population"), lwd=4)

pairs(cbind(region1$LogY_i, log(region1$total_population)), labels = c("Total serious crimes (log-scale)", "Total population (log-scale)"), lwd = 4)
```

## Region 2

```{r}
# log(Y_i) = Beta_0 + Beta_1*total_population

boxplot(cbind(region1$LogY_i, region1$Physicians),names = c("Total serious crimes (log-scale)", "Physicians"))

pairs(cbind(region1$LogY_i, region1$Physicians),labels=c("Total serious crimes  (log-scale)", "Physicians"), lwd=4)

pairs(cbind(region1$LogY_i, log(region1$Physicians)), labels = c("Total serious crimes (log-scale)", "Physicians (log-scale)"), lwd = 4)
```


## Region 3

```{r}
# log(Y_i) = Beta_0 + Beta_1*total_population

boxplot(cbind(region1$LogY_i, region1$personal_income),names = c("Total serious crimes (log-scale)", "Personal Income"))

pairs(cbind(region1$LogY_i, region1$personal_income),labels=c("Total serious crimes  (log-scale)", "Personal Income"), lwd=4)

pairs(cbind(region1$LogY_i, log(region1$personal_income)), labels = c("Total serious crimes (log-scale)", "Personal Income (log-scale)"), lwd = 4)
```


## Region 4

```{r}
# Region 1
# log(Y_i) = Beta_0 + Beta_1*total_population

boxplot(cbind(region1$LogY_i, region1$Population_18to34),names = c("Total serious crimes (log-scale)", "Population (Age 18-34)"))

pairs(cbind(region1$LogY_i, region1$Population_18to34),labels=c("Total serious crimes  (log-scale)", "Population (Age 18-34)"), lwd=4)

pairs(cbind(region1$LogY_i, log(region1$Population_18to34)), labels = c("Total serious crimes (log-scale)", "Population (Age 18-34) (log-scale)"), lwd = 4)
```


```{r}
# Total Population
region1 |>
        ggplot(aes(x = total_population, y = LogY_i)) +
        geom_point() +
        theme_bw()

# Physicians
region2 |>
        ggplot(aes(x = Physicians)) +
        geom_histogram() +
        theme_bw()

region2 |>
        ggplot(aes(x = Physicians, y = LogY_i)) +
        geom_point() +
        theme_bw()

# Personal Income
region3 |>
        ggplot(aes(x = personal_income, y = LogY_i)) +
        geom_point() +
        theme_bw()

# Population 18-34
region4 |>
        ggplot(aes(x = Population_18to34, y = LogY_i)) +
        geom_point() +
        theme_bw()
```

CHOSEN PREDICTORS for SLR (no repeats):

Region 1: Total Population

Region 2: Physicians

Region 3: Personal Income

Region 4: Population 18-34

```{r SLR models}
model_1 <- lm(LogY_i ~ total_population, data = region1)
model_2 <- lm(LogY_i ~ Physicians, data = region2)
model_3 <- lm(LogY_i ~ personal_income, data = region3)
model_4 <- lm(LogY_i ~ Population_18to34, data = region4)
```

# Question 2 : Model Evaluation

```{r model stats & CI}
# REGION 1
summary(model_1) # all model stats
plot(model_1) # all four diagnostic plots
confint(model_1) # 95% CI for both b0 and b1

# REGION 2
summary(model_2)
plot(model_2) # insane outlier here
confint(model_2)

# REGION 3
summary(model_3)
plot(model_3) # & here
confint(model_3)

# REGION 4
summary(model_4)
plot(model_4)
confint(model_4)
```

**REGION 1**

For Region 1, we fit a simple linear regression of log(total serious crimes) on the estimated total population. The estimated slope is 0.000002526 (p = \< 2.2e-16 or \~0), suggesting a positive and significant relationship. The number is small despite the relationship being strong because total population is very large. The model explains 74.15% of the variation in log(serious crimes).

The residual plots show a fanning pattern suggesting non-constant variance (heteroscedasticity), which motivates the transformation approach explored in Question 3. Q-Q looks good, residuals appear approximately normally distributed. There are multiple influential points pointed out in the Residuals v. Leverage plot (points 6, 11, & 16). Assumptions are not reasonably met.

We are 95% confident the intercept is between [8.09, 8.40] and that for each one unit increase in total population, log(Y) increases by between [0.0000022, 0.0000028]. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant.

**REGION 2**

For Region 2, we fit a simple linear regression of log(total serious crimes) on the number of professionally active non-federal physicians. The estimated slope is 8.861 (p = 7.134e-15 or \~0), suggesting a positive and significant relationship. The model explains 43.14% of the variation in log(serious crimes).

The residual plots show extreme fanning suggesting heteroscedasticity again. Points veer away from the ends of the Q-Q plot line, so residuals are not entirely normally distributed. There is an extremely influential outlier (point 2) pointed out in the Residuals v. Leverage plot that greatly exceeds Cook's distance and is likely pulling hard on the regression line. Assumptions are not reasonably met.

We are 95% confident the intercept is between [8.67, 9.04] and that for each physician, log(Y) increases by between [0.0003, 0.0005]. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant.

**REGION 3**

For Region 3, we fit a simple linear regression of log(total serious crimes) on the total personal income of CDI population (in millions of dollars). The estimated slope is 0.0001013 (p = \< 2.2e-16 or \~0), suggesting a positive and significant relationship. The model explains 61.83% of the variation in log(serious crimes).

Once again, residual plots show extreme fanning suggesting heteroscedasticity. Points veer only slightly away from Q-Q plot, residuals appear approximately normally distributed. Again, there is an extremely influential outlier (point 3) that greatly exceeds Cook's distance. Assumptions are not reasonably met.

We are 95% confident the intercept is between [8.93, 9.17] and that for each one unit increase in personal income, log(Y) increases by between [0.0000885, 0.0001]. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant.

**REGION 4**

For Region 4, we fit a simple linear regression of log(total serious crimes) on percent of CDI population aged 18-34. The estimated slope is 0.1495 (p = 3.248e-05, quite small), suggesting a positive and significant relationship. The model explains 19.63% of the variation in log(serious crimes), which is quite low! This predictor is significant but not strong.

Residual plot is approximately randomly distributed, suggesting mostly constant variance. Q-Q points closely follow line. Some influential points (points 433 and 325) come close to Cook's distance cutoff but don't exceed it. Assumptions are reasonably met!

We are 95% confident the intercept is between [3.75, 7.57] and that for each one unit increase in population of 18-34 year olds, log(Y) increases by between [0.082, 0.21]. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant.

# Question 3 : Model Refinement Using Transformations

```{r}

```
