---
title: "STA 108 Final Project"
author: "Jachin Lee, Jacob Melendrez, Nadia Bourne, Olivia Ma, and Tarilyn Tong"
output:
  pdf_document: default
  html_notebook: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.width = 4, fig.height = 4)
```


```{r link cloned repo so you can pull/commit/push}
# RUN THESE ALL ONCE!!! un-comment the code blocks as you go
#install.packages(c("usethis", "gitcreds"))

# then run this:
#usethis::create_github_token()

# which opens GitHub in your browser, generate a token
# COPY it before closing the page

# run this and paste token into terminal when prompted
#gitcreds::gitcreds_set()

# done! :) make sure to re-comment the code blocks
```

```{r libaries, include=FALSE}
# make sure to download any you dont have with: install.packages("package_name")
library(here)
library(corrplot)
library(tidyverse)
```

# Question 1 : Exploratory Analysis and Model Specification

The Demographic data set was first explored using descriptive statistics and correlation analyses in order to understand the relationships between crime rates and demographic variables across four regions. The data set was split into four regions, region 1 containing 103 observations, region 2 containing 108 observations, region 3 containing 152 observations and region 4 containing 77 observations. Summary statistics were used to explore and evaluate the ranges, distributions, and variability  of the demographic and socioeconomic variables.

The correlation matrices and plots were used to examine the relationships between the variables. If a variable contained a stronger correlation to the response variable, they were selected as potential predictors for the regional regression models. Separate analyses was done for each region to determine whether different predictors were associated with crime rates. This analysis suggested that different demographic variables were more strongly associated with crime depending on the region. In this project, we selected total population, the number of physicians, personal income, and population of ages 18 to 34.

```{r prep}
Demographic <- read.table(here("data", "Demographic.txt")) # more robust read-in

# change column names to match correct variables
colnames(Demographic) <- c('ID','County','state','Land_area','total_population',
                          'Population_18to34','Population_65','Physicians','beds',
                          'Y_i','Graduate_highschool','Graduate_Bachelor','below_poverty',
                          'unemployment','capita_income','personal_income','Geographic_region')
Demographic2 <- Demographic

# Variable names and dimensions of the table
head(Demographic2)
names(Demographic2)
dim(Demographic2)
summary(Demographic2)

Demographic2$Y_i <- log(Demographic$Y_i) # log transform Y

colnames(Demographic2)[10] <- "LogY_i"

region1 <- Demographic2[Demographic2$Geographic_region == 1,]
region2 <- Demographic2[Demographic2$Geographic_region == 2,]
region3 <- Demographic2[Demographic2$Geographic_region == 3,]
region4 <- Demographic2[Demographic2$Geographic_region == 4,]

n1<-nrow(region1)
n2<-nrow(region2)
n3<-nrow(region3)
n4<-nrow(region4)
```

```{r corr plots, warning=FALSE}
# general corr
cor(Demographic2[,4:17])
corrplot(cor(Demographic2[,4:17]))

# per region corr
regions <- list(region1, region2, region3, region4)
top_list <- list()

for (i in seq_along(regions)) {
  
  cat("FOR REGION: ", i, "\n")
  
  region <- regions[[i]]
  cors <- sapply(region[, 4:17], function(x) cor(x, region$LogY_i))
  ranked <- sort(abs(cors), decreasing = TRUE)
  
  top_list[[i]] <- head(ranked, 5)
  
  corrplot(cor(region$LogY_i, region[,4:17]))
  
  print(top_list[[i]])
}
```

## Region 1

Region 1 consists of total population as the predictor. A boxplot was created and indicated variability within county population size, while the scatter plots suggest that there is a positive relationship between total population and the log-transformed response variability, crime rates. The additional scatter plot displays the log-transformed total population and indicates whether there is improved linearity of the relationship.

```{r}
# log(Y_i) = Beta_0 + Beta_1*total_population

boxplot(cbind(region1$LogY_i, region1$total_population),names = c("Total 
                                                                  serious crimes (log-scale)", "Total Population"))

pairs(cbind(region1$LogY_i, region1$total_population),labels = c("Total 
                                                               serious crimes (log-scale)", "Total Population"), lwd=4)

pairs(cbind(region1$LogY_i, log(region1$total_population)), labels = c("Total 
                                                                       serious crimes (log-scale)", "Total population (log-scale)"), lwd = 4)
```
Here we can see that when comparing the two scatter plots, the transformed predictor (total population) has reduced skewness and produces a more consistent linear trend.

## Region 2

```{r}
# log(Y_i) = Beta_0 + Beta_1*total_population

boxplot(cbind(region1$LogY_i, region1$Physicians),names = c("Total serious 
      crimes (log-scale)", "Physicians"))

pairs(cbind(region1$LogY_i, region1$Physicians),labels=c("Total serious crimes 
      (log-scale)", "Physicians"), lwd=4)

pairs(cbind(region1$LogY_i, log(region1$Physicians)), labels = c("Total serious 
      crimes (log-scale)", "Physicians (log-scale)"), lwd = 4)
```

Region 2 displays the number of physicians as a predictor variable. The plots indicate that the counties with higher numbers of physicians generally have higher levels of crime after transformation. This transformed relationship appears more linear and less influenced by extreme observations.

## Region 3

Personal income was chosen as the predictor variable. The boxplot shows variation in income levels across counties, and the scatter-plots suggested an association between personal income and the transformed crime variable. The logged transformation of personal income was evaluated to determine whether it improved the distributional properties and linear relationship with the response variable.

```{r}
# log(Y_i) = Beta_0 + Beta_1*total_population

boxplot(cbind(region1$LogY_i, region1$personal_income),names = c("Total serious crimes (log-scale)", "Personal Income"))

pairs(cbind(region1$LogY_i, region1$personal_income),labels = c("Total serious crimes  (log-scale)", "Personal Income"), lwd=4)

pairs(cbind(region1$LogY_i, log(region1$personal_income)), labels = c("Total serious crimes (log-scale)", "Personal Income (log-scale)"), lwd = 4)
```


## Region 4

For Region 4, the population between the ages of 18 and 34 was selected as the predictor variable. The exploratory analysis suggested a positive association between the younger adult population and the transformed crime variable. Additional scatter-plots using the log of the predictor were used to assess whether the transformation improved linearity and reduced the effect of skewness.

```{r}
# Region 1
# log(Y_i) = Beta_0 + Beta_1*total_population

boxplot(cbind(region1$LogY_i, region1$Population_18to34),names = c("Total serious crimes (log-scale)", "Population (Age 18-34)"))

pairs(cbind(region1$LogY_i, region1$Population_18to34),labels=c("Total serious crimes  (log-scale)", "Population (Age 18-34)"), lwd=4)

pairs(cbind(region1$LogY_i, log(region1$Population_18to34)), labels = c("Total serious crimes (log-scale)", "Population (Age 18-34) (log-scale)"), lwd = 4)
```


### CHOSEN PREDICTORS for SLR (no repeats):

- Region 1: Total Population

- Region 2: Physicians

- Region 3: Personal Income

- Region 4: Population 18-34

```{r SLR models}
model_1 <- lm(LogY_i ~ total_population, data = region1)
model_2 <- lm(LogY_i ~ Physicians, data = region2)
model_3 <- lm(LogY_i ~ personal_income, data = region3)
model_4 <- lm(LogY_i ~ Population_18to34, data = region4)
```

# Question 2 : Model Evaluation

```{r model stats & CI}
# REGION 1
summary(model_1) # all model stats
plot(model_1) # all four diagnostic plots
confint(model_1) # 95% CI for both b0 and b1

# REGION 2
summary(model_2)
plot(model_2) # insane outlier here
confint(model_2)

# REGION 3
summary(model_3)
plot(model_3) # & here
confint(model_3)

# REGION 4
summary(model_4)
plot(model_4)
confint(model_4)
```

**REGION 1**

For Region 1, we fit a simple linear regression of log(total serious crimes) on the estimated total population. The estimated slope is 0.000002526 (p = \< 2.2e-16 or \~0), suggesting a positive and significant relationship. The number is small despite the relationship being strong because total population is very large. The model explains 74.15% of the variation in log(serious crimes).

The residual plots show a fanning pattern suggesting non-constant variance (heteroscedasticity), which motivates the transformation approach explored in Question 3. Q-Q looks good, residuals appear approximately normally distributed. There are multiple influential points pointed out in the Residuals v. Leverage plot (points 6, 11, & 16). Assumptions are not reasonably met.

We are 95% confident the intercept is between [8.09, 8.40] and that for each one unit increase in total population, log(Y) increases by between [0.0000022, 0.0000028]. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant.

**REGION 2**

For Region 2, we fit a simple linear regression of log(total serious crimes) on the number of professionally active non-federal physicians. The estimated slope is 0.0004375 (p = 7.134e-15 or \~0), suggesting a positive and significant relationship. The model explains 43.14% of the variation in log(serious crimes).

The residual plots show extreme fanning suggesting heteroscedasticity again. Points veer away from the ends of the Q-Q plot line, so residuals are not entirely normally distributed. There is an extremely influential outlier (point 2) pointed out in the Residuals v. Leverage plot that greatly exceeds Cook's distance and is likely pulling hard on the regression line. Assumptions are not reasonably met.

We are 95% confident the intercept is between [8.67, 9.04] and that for each physician, log(Y) increases by between [0.0003, 0.0005]. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant.

**REGION 3**

For Region 3, we fit a simple linear regression of log(total serious crimes) on the total personal income of CDI population (in millions of dollars). The estimated slope is 0.0001013 (p = \< 2.2e-16 or \~0), suggesting a positive and significant relationship. The model explains 61.83% of the variation in log(serious crimes).

Once again, residual plots show extreme fanning suggesting heteroscedasticity. Points veer only slightly away from Q-Q plot, residuals appear approximately normally distributed. Again, there is an extremely influential outlier (point 3) that greatly exceeds Cook's distance. Assumptions are not reasonably met.

We are 95% confident the intercept is between [8.93, 9.17] and that for each one unit increase in personal income, log(Y) increases by between [0.000088, 0.00011]. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant.

**REGION 4**

For Region 4, we fit a simple linear regression of log(total serious crimes) on percent of CDI population aged 18-34. The estimated slope is 0.1495 (p = 3.248e-05, quite small), suggesting a positive and significant relationship. The model explains 19.63% of the variation in log(serious crimes), which is quite low! This predictor is significant but not strong.

Residual plot is approximately randomly distributed, suggesting mostly constant variance. Q-Q points closely follow line. Some influential points (points 433 and 325) come close to Cook's distance cutoff but don't exceed it. Assumptions are reasonably met!

We are 95% confident the intercept is between [3.75, 7.57] and that for each one unit increase in population of 18-34 year olds, log(Y) increases by between [0.082, 0.21]. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant.

# Question 3 : Model Refinement Using Transformations

```{r}
# REGION 1
model_1b <- lm(log(Demographic$Y_i) ~ log(Demographic$total_population), data = region1)

anova(model_1b)
summary(model_1b)
plot(model_1b) # greatly improved linearity
confint(model_1b)

# REGION 2
model_2b <- lm(log(Demographic$Y_i) ~ log(Demographic$Physicians), data=region2)

anova(model_2b)
summary(model_2b)
plot(model_2b)
confint(model_2b)

# REGION 3
model_3b <- lm(log(Demographic$Y_i) ~ log(Demographic$personal_income), data = region3)

anova(model_3b)
summary(model_3b)
plot(model_3b) # greatly improved linearity, previous outliers arent as bad now
confint(model_3b)

# REGION 4
model_4b <- lm(log(Demographic$Y_i) ~ log(Demographic$Population_18to34), data = region4)

anova(model_4b)
summary(model_4b)
plot(model_4b)
confint(model_4b)
```

**REGION 1**

For Region 1, we applied a log transformation onto the predictor (estimated total population). The estimated slope is 1.234, suggesting a positive and significant relationship. The model now explains 81.27% (R^2 = 0.8127) of the variation in log(serious crimes), an increase from the previous 74.15%.

Compared to the original model, the residuals plots of the transformed model display a drastic increase in homoscedastic residuals. However, when plotting the residuals, the data points appear to follow a slight funnel shape, though this does not appear to be dramatic enough to constitute an egregious violation of the assumption of constant variance. On the Q-Q plot, the points still mostly form a straight line, but has some points veering off near the ends. Compared to the original model, it appears that the transformation may have decreased normality. However, it is only a few points veering off, with the vast majority forming a straight line, suggesting that this does not egregiously violate the assumption of normality. In regards to the Residuals vs. Leverage plot, the previously influential points (points 6, 11, & 16) of the original model are no longer as influential in the transformed model. Assumptions are therefore now reasonably met. Despite the apparent decrease in normality, the transformation reduced the impact of influential points and radically increased homoscedasticity in the residuals. The model therefore overall appears to be a better fit for the data after the transformation and the quality of inference seems to have dramatically improved for these reasons.

We are 95% confident the intercept is between [-6.594, -5.204] and that for each one unit increase in personal income, log(Y) increases by between [1.178, 1.290]. The confidence intervals for both beta_0 and beta_1 became much wider. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant. The p-values for beta_0 and beta_1 remained the same at 2e^-16 after the transformation.

**REGION 2**

For Region 2, we applied the log transformation to the original model, the slope of the model is approximately 0.806 (p = <2e^-16 or \~0), suggesting a positive and significant relationship. The p-value for the transformed model is slightly smaller than the original model however both are approximately zero. The transformed CI [0.760, 0.853] is not smaller than the original model, while this initially may indicate that the original model is better. However with further analysis of the Residuals vs. Fitted we can see significant improvement with the transformed data where the data is showing homoscedasticity in the Residuals vs. Fitted plot. We can also see in the Residuals vs. Leverage plot that the mentioned outlier is no longer present and is better fitting. The Q-Q Residuals plot also shows the residuals being approximately normally distributed but not much variation from the original model in this plot. The R^2 value is higher for the transformed data original (R^2 = 0.4314), transformed (R^2 = 0.7267), we can see a large increase in the model variability being explained by the predictor variable which is physicians in this case. Based off the improvement in the data matching the transformed model we will choose the log transformation to represent our model for region 2. 


**REGION 3**

For Region 3, we applied a log transformation onto the predictor (personal income of CDI population (in millions of dollars)). The estimated slope is 1.035, suggesting a positive and significant relationship. The model now explains 73.7% (R^2 = 0.737) of the variation in log(serious crimes), an increase from the previous 61.83%.

Contrary to the original model, the residuals plots of the transformed model display homoscedasticity. On the Q-Q plot, the points still mostly form a straight line, with only a few points veering off slightly near the ends. Therefore, the residuals appear approximately normally distributed just as it was with the original model. The previous outlier in the original model (point 3) no longer exceeds Cook's Distance, suggesting a decrease in its influence. Assumptions are now reasonably met. The model overall appears to be a better fit for the data after the transformation and the quality of inference seems to have dramatically improved for these reasons.

We are 95% confident the intercept is between [0.242, 1.23] and that for each one unit increase in personal income, log(Y) increases by between [0.977, 1.093]. The confidence intervals for both beta_0 and beta_1 became much wider. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant. The p-value for beta_0 increased from 2e-16 to 0.00356, while the p-value for beta_1 remained the same at 2.2e-16 after the transformation.

**REGION 4**
For Region 4, we applied the log transformation to the original model, the slope of the model is approximately 1.343 (p = 2.26e^-04), suggesting a positive and significant relationship. The p-value for the transformed model is larger than the original model (p = 3.25e^-05) but both models indicate the predictor variable being significant. The transformed CI [0.633, 2.053] is much larger than the original model CI [0.082, 0.21], this indicates that the transformed model may not be a better model than the original. When looking at both the Residuals vs. Fitted and Q-Q plots we can see that both the original and transformed models are very similar with slight variation but by looking at each plot we can see homoscedasticity and the residuals being normally distributed. So the transformed model did not explicitly improve the fit of the data compared to the original. Looking at the R^2 values, original (R^2 = 0.1963) and transformed (R^2 = 0.0284), we can see that the transformed model is worse at explaining the variation being caused by the predictor variable which is Population 18 to 34. Based off the values and plots, we will not choose the log transformated model to represent region 4 but stick with the original as the log transformation did not improve the model's ability to explain the relationship between total number of crimes and Population 18 to 34. 

```{r}
# Total Population
region1 |>
        ggplot(aes(x = log(total_population), y = LogY_i)) +
        geom_point() +
        theme_bw()

# Physicians
region2 |>
        ggplot(aes(x = log(Physicians))) +
        geom_histogram() +
        theme_bw()

region2 |>
        ggplot(aes(x = log(Physicians), y = LogY_i)) +
        geom_point() +
        theme_bw()

# Personal Income
region3 |>
        ggplot(aes(x = log(personal_income), y = LogY_i)) +
        geom_point() +
        theme_bw()

# Population 18-34
region4 |>
        ggplot(aes(x = log(Population_18to34), y = LogY_i)) +
        geom_point() +
        theme_bw()
```

# Question 4

## 4.1 Multiple Linear Regression Analysis
```{r}
# Setting up data and choosing a region

nrow(region4)
head(region4)

pairs(region4[,5:12], lwd=2)
cor(region4[,5:12])
```

```{r}
model_TP <- lm(LogY_i ~ total_population, data = region4)
model_logTP <- lm(LogY_i ~ log(total_population), data = region4)

plot(model_TP)
summary(model_TP)
plot(model_logTP)
summary(model_logTP)

model_P <- lm(LogY_i ~ Physicians, data = region4)
model_logP <- lm(LogY_i ~ log(Physicians), data = region4)

plot(model_P)
summary(model_P)
plot(model_logP)
summary(model_logP)

model_PI <- lm(LogY_i ~ personal_income, data = region4)
model_logPI <- lm(LogY_i ~ log(personal_income), data = region4)

plot(model_PI)
summary(model_PI)
plot(model_logPI)
summary(model_logPI)
```

**POPULATION 18-34 INTERPRETATION**

This is the predictor originally used in Questions 1-3. For the same reasons mentioned in Q3, transformations did not improve this predictor, leaving the untransformed model as the best fit.

**TOTAL POPULATION INTERPRETATION**

A log transformation was the most appropriate for this predictor (total population). We see immediately from the R^2 values that the log transformation nearly doubled the R^2 value (0.4868 untransformed, 0.8741 transformed), suggesting that the transformation much better explains the variance of the data and therefore results in a better fit. We can also see from the Residual vs. Fitted plots that the transformation greatly increased heteroscedasticity, as the untransformed plot demonstrates a curved pattern in the residuals while the transformed plot displays much more randomness in the residuals. The influence of the outlier in the untransformed Q-Q plot is also greatly reduced after the transformation, in which the outlier point fits much better with the rest of the points. This is also supported by the Residuals vs. Leverage plot, in which the point that originally exceeds Cook's Distance in the untransformed plot is now well within the distance after the transformation. Other transformations, such as a square root transformation or a square transformation, resulted in lower R^2 values, more homoscedasticity, and reduced normality compared to the log transformation, suggesting that the log transformation was the best transformation on this predictor in maintaining constant variance and normality.

**PHYSICIANS INTERPRETATION**

A log transformation was the most appropriate for this predictor (number of physicians). We see immediately from the R^2 values that the log transformation nearly doubled the R^2 value (0.5011 untransformed, 0.9083 transformed), suggesting that the transformation much better explains the variance of the data and therefore results in a better fit. We can also see from the Residual vs. Fitted plots that the transformation greatly increased heteroscedasticity, as the untransformed plot demonstrates a curved pattern in the residuals while the transformed plot displays much more randomness in the residuals. The influence of the outlier in the untransformed Q-Q plot is also greatly reduced after the transformation, in which the outlier point fits much better with the rest of the points. This is also supported by the Residuals vs. Leverage plot, in which the point that originally exceeds Cook's Distance in the untransformed plot is now well within the distance after the transformation. Other transformations, such as a square root transformation or a square transformation, resulted in lower R^2 values, more homoscedasticity, and reduced normality compared to the log transformation, suggesting that the log transformation was the best transformation on this predictor in maintaining constant variance and normality.

**PERSONAL INCOME INTERPRETATION**

A log transformation was the most appropriate for this predictor (personal income). We see immediately from the R^2 values that the log transformation nearly doubled the R^2 value (0.507 untransformed, 0.9537 transformed), suggesting that the transformation much better explains the variance of the data and therefore results in a better fit. We can also see from the Residual vs. Fitted plots that the transformation greatly increased heteroscedasticity, as the untransformed plot demonstrates a curved pattern in the residuals while the transformed plot displays much more randomness in the residuals. The influence of the outlier in the untransformed Q-Q plot is also greatly reduced after the transformation, in which the outlier point fits much better with the rest of the points. This is also supported by the Residuals vs. Leverage plot, in which the point that originally exceeds Cook's Distance in the untransformed plot is now well within the distance after the transformation. Other transformations, such as a square root transformation or a square transformation, resulted in lower R^2 values, more homoscedasticity, and reduced normality compared to the log transformation, suggesting that the log transformation was the best transformation on this predictor in maintaining constant variance and normality.

```{r}
model <- lm(LogY_i ~ Population_18to34 + log(total_population) + log(Physicians) + log(personal_income), data = region4)

plot(model)
summary(model)
anova(model)
```

**INTERPRETATION**




## 4.2 Model Comparison Using 2p−1 Models
```{r}
# log transformations turned out to be the best for the extra 3 predictors
model_4_1 <- lm(LogY_i ~ 1, data = region4)

model_4_2 <- lm(LogY_i ~ Population_18to34, data = region4)

model_4_3 <- lm(LogY_i ~ log(total_population), data = region4)

model_4_4 <- lm(LogY_i ~ log(Physicians), data = region4)

model_4_5 <- lm(LogY_i ~ log(personal_income), data = region4)

model_4_6 <- lm(LogY_i ~ Population_18to34 + log(total_population), data = region4)

model_4_7 <- lm(LogY_i ~ Population_18to34 + log(Physicians), data = region4)

model_4_8 <- lm(LogY_i ~ Population_18to34 + log(personal_income), data = region4)

model_4_9 <- lm(LogY_i ~ log(total_population) + log(Physicians), data = region4)

model_4_10 <- lm(LogY_i ~ log(total_population) + log(personal_income), data = region4)

model_4_11 <- lm(LogY_i ~ log(Physicians) + log(personal_income), data = region4)

model_4_12 <- lm(LogY_i ~ Population_18to34 + log(total_population) + log(Physicians), data = region4)

model_4_13 <- lm(LogY_i ~ Population_18to34 + log(total_population) + log(personal_income), data = region4)

model_4_14 <- lm(LogY_i ~ Population_18to34 + log(Physicians) + log(personal_income), data = region4)

model_4_15 <- lm(LogY_i ~ log(total_population) + log(Physicians) + log(personal_income), data = region4)

model_4_16 <- lm(LogY_i ~ Population_18to34 + log(total_population) + log(Physicians) + log(personal_income), data = region4)

models <- list(model_4_1, model_4_2, model_4_3, model_4_4, model_4_5, model_4_6, model_4_7, model_4_8, model_4_9, model_4_10, model_4_11, model_4_12, model_4_13, model_4_14, model_4_15, model_4_16)
```

```{r}
n <- nrow(region4)
p_full <- length(coef(model_4_16))
MSE_full <- sum(residuals(model_4_16)^2) / (n - p_full)

results <- data.frame(
  variables = c("None", "Population_18to34", "log(Total_Population)", "log(Physicians)", "log(Personal_Income)", "18to34+logTotPop", "18to34+logPhys", "18to34+logIncome", "logTotPop+logPhys", "logTotPop+logIncome", "logPhys+logIncome", "18to34+logTotPop+logPhys", "18to34+logTotPop+logIncome", "18to34+logPhys+logIncome", "logTotPop+logPhys+logIncome", "All four"),
  p = sapply(models, function(m) length(coef(m))),
  SSE = sapply(models, function(m) sum(residuals(m)^2)),
  R2 = sapply(models, function(m) summary(m)$r.squared),
  R2_adj = sapply(models, function(m) summary(m)$adj.r.squared),
  C_p = sapply(models, function(m) {
               SSE_p <- sum(residuals(m)^2)
               p_p   <- length(coef(m))
               SSE_p / MSE_full - (n - 2 * p_p)
               }),
  AIC = sapply(models, AIC),
  BIC = sapply(models, BIC),
  PRESS = sapply(models, function(m) {
                 sum((residuals(m) / (1 - hatvalues(m)))^2)
                 }),
  F_pval = sapply(models, function(m) {
                  f <- summary(m)$fstatistic
                  if (is.null(f)) return(NA)  # null model has no F stat
                  pf(f[1], f[2], f[3], lower.tail = FALSE)
                  }),
  sig_predictors = sapply(models, function(m) {
                          coefs <- summary(m)$coefficients
                          sum(coefs[-1, 4] < 0.05) # exclude intercept row, count p < 0.05
                          })
)

results
```
**MODEL COMPARISON INTERPRETATION**

A series of 16 multiple linear regression models were fitted to explain variation in the log-transformed response variable log(Yi). The models include different combinations of the four predictors. The table shows a clear improvement in model fit when moving the intercept model to models that include predictors. For example, the null model explains none of the variation (R^2 = 0, SSE = 88.01), and adds a single predictor that improves fitness, such as log(Total_Population) achieves R^2 = 0.954, reducing the SSE to about 4.08. This concludes that with additional predictors, improvements become smaller indicating diminishing returns.
