---
title: "project!!!!!"
output: html_notebook
---

```{r link cloned repo so you can pull/commit/push}
# RUN THESE ALL ONCE!!! un-comment the code blocks as you go
#install.packages(c("usethis", "gitcreds"))

# then run this:
#usethis::create_github_token()

# which opens GitHub in your browser, generate a token
# COPY it before closing the page

# run this and paste token into terminal when prompted
#gitcreds::gitcreds_set()

# done! :) make sure to re-comment the code blocks
```

```{r libaries, include=FALSE}
# make sure to download any you dont have with: install.packages("package_name")
library(here)
library(corrplot)
library(tidyverse)
```

# Question 1 : Exploratory Analysis and Model Specification

```{r prep}
Demographic <- read.table(here("data", "Demographic.txt")) # more robust read-in

# change column names to match correct variables
colnames(Demographic) <- c('ID','County','state','Land_area','total_population',
                          'Population_18to34','Population_65','Physicians','beds',
                          'Y_i','Graduate_highschool','Graduate_Bachelor','below_poverty',
                          'unemployment','capita_income','personal_income','Geographic_region')
Demographic2 <- Demographic

# Variable names and dimensions of the table
head(Demographic2)
names(Demographic2)
dim(Demographic2)
summary(Demographic2)

Demographic2$Y_i <- log(Demographic$Y_i) # log transform Y

colnames(Demographic2)[10] <- "LogY_i"

region1 <- Demographic2[Demographic2$Geographic_region == 1,]
region2 <- Demographic2[Demographic2$Geographic_region == 2,]
region3 <- Demographic2[Demographic2$Geographic_region == 3,]
region4 <- Demographic2[Demographic2$Geographic_region == 4,]

n1<-nrow(region1)
n2<-nrow(region2)
n3<-nrow(region3)
n4<-nrow(region4)
```

```{r corr plots, warning=FALSE}
# general corr
cor(Demographic2[,4:17])
corrplot(cor(Demographic2[,4:17]))

# per region corr
regions <- list(region1, region2, region3, region4)
top_list <- list()

for (i in seq_along(regions)) {
  
  cat("FOR REGION: ", i, "\n")
  
  region <- regions[[i]]
  cors <- sapply(region[, 4:17], function(x) cor(x, region$LogY_i))
  ranked <- sort(abs(cors), decreasing = TRUE)
  
  top_list[[i]] <- head(ranked, 5)
  
  corrplot(cor(region$LogY_i, region[,4:17]))
  
  print(top_list[[i]])
}
```

```{r}
# Region 1
# log(Y_i) = Beta_0 + Beta_1*total_population

boxplot(cbind(region1$LogY_i,region1$below_poverty, region1$unemployment),names = c("Total serious crimes (log-scale)", "Poverty Levels", "Percent Unemployment"))

pairs(cbind(region1$LogY_i, region1$below_poverty, region1$total_population), labels = c("Total serious crimes", "Poverty Level", "Total population"), lwd = 4)

pairs(cbind(region1$LogY_i, region1$below_poverty, region1$total_population),labels=c("Total serious crimes  (log-scale)","Poverty level ", "Poverty Levels", "Total Population"), lwd=4)

pairs(cbind(region1$LogY_i, region1$below_poverty, log(region1$total_population)), labels = c("Total serious crimes (log-scale)","Poverty Levels", "Total population (log-scale)"), lwd = 4)
```

```{r}
# Total Population
region_1 |>
        ggplot(aes(x = total_population, y = LogY_i)) +
        geom_point() +
        theme_bw()

# Physicians
region_2 |>
        ggplot(aes(x = Physicians)) +
        geom_histogram() +
        theme_bw()

region_2 |>
        ggplot(aes(x = Physicians, y = LogY_i)) +
        geom_point() +
        theme_bw()

# Personal Income
region_3 |>
        ggplot(aes(x = personal_income, y = LogY_i)) +
        geom_point() +
        theme_bw()

# Population 18-34
region_4 |>
        ggplot(aes(x = Population_18to34, y = LogY_i)) +
        geom_point() +
        theme_bw()
```

CHOSEN PREDICTORS for SLR (no repeats):

Region 1: Total Population

Region 2: Physicians

Region 3: Personal Income

Region 4: Population 18-34

```{r SLR models}
model_1 <- lm(LogY_i ~ total_population, data = region_1)
model_2 <- lm(LogY_i ~ Physicians, data = region_2)
model_3 <- lm(LogY_i ~ personal_income, data = region_3)
model_4 <- lm(LogY_i ~ Population_18to34, data = region_4)
```

# Question 2 : Model Evaluation

```{r model stats & CI}
# REGION 1
summary(model_1) # all model stats
plot(model_1) # all four diagnostic plots
confint(model_1) # 95% CI for both b0 and b1

# REGION 2
summary(model_2)
plot(model_2) # insane outlier here
confint(model_2)

# REGION 3
summary(model_3)
plot(model_3) # & here
confint(model_3)

# REGION 4
summary(model_4)
plot(model_4)
confint(model_4)
```

GENERIC TEXT TO REUSE: For Region [X], we fit a simple linear regression of log(total serious crimes) on [predictor]. The estimated slope is [β₁] (p = ...), suggesting a [positive/negative] and [significant/not significant] relationship. The model explains [R²]% of the variation in log(serious crimes). 

**REGION 1**

For Region 1, we fit a simple linear regression of log(total serious crimes) on the total population. The estimated slope is 0.000002526 (p = < 2.2e-16 or \~0), suggesting a positive and significant relationship. The number is small despite the relationship being strong because total population is very large. The model explains 74.15% of the variation in log(serious crimes). 

The residual plots show a fanning pattern suggesting non-constant variance \(heteroscedasticity\), which motivates the transformation approach explored in Question 3. Q-Q looks good, residuals appear approximately normally distributed. There are multiple influential points pointed out in the Residuals v. Leverage plot (points 6, 11, & 16). 

We are 95% confident the intercept is between [8.09, 8.40] and that for each one unit increase in total population, log(Y) increases by between [0.0000022, 0.0000028]. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant.

**REGION 2**

For Region 2, we fit a simple linear regression of log(total serious crimes) on the number of physicians. The estimated slope is 8.861 (p = 7.134e-15 or \~0), suggesting a positive and significant relationship. The model explains 43.14% of the variation in log(serious crimes). 

The residual plots show extreme fanning suggesting heteroscedasticity again. Points veer away from the ends of the Q-Q plot line, so residuals are not entirely normally distributed. There is an extremely influential outlier (point 2) pointed out in the Residuals v. Leverage plot that greatly exceeds Cook's distance.

We are 95% confident the intercept is between [8.67, 9.04] and that for each physician, log(Y) increases by between [0.0003, 0.0005]. Notably, the CI for beta_1 does not include 0, which is evidence the predictor is significant.

**REGION 3**

Summary statistics:

Diagnostic plots:

CI Interpretation: We are 95% confident the intercept is between [lwr, upr] and that for each one unit increase in personal income, log(Y) increases by between [, ].

**REGION 4**

Summary statistics:

Diagnostic plots:

CI Interpretation: We are 95% confident the intercept is between [lwr, upr] and that for each one unit increase in total population of 18-34 year olds, log(Y) increases by between [, ].

# Question 3 : Model Refinement Using Transformations

```{r}

```
